# Reproducible Research: Peer Assessment 1
<!-- 
Load required R libraries
-->
First we'll start by loading all the needed libraries
```{r}
library(lattice)
library(ggplot2)
library(lubridate)
library(xtable)
library(plyr)
```

<!--
modify
-->

## Loading and preprocessing the data
We will start by loading and doing some processing of the dataset we have

```{r loaddata}
data <- read.table(unz("./activity.zip", "activity.csv"), header=TRUE, sep=",", na.strings="NA")
```

Take a quick look at the data with the `str` command in R: 

```{r}
str(data)
```

Next we will convert the `date` column in the dataset to date values with the `lubridate` package

```{r processdata}
data$date <- ymd(data$date)
```

And our new data frame looks like: 

```{r}
str(data)
```

## What is mean total number of steps taken per day?

First let's make a histogram of the total number of steps taken per day.   
Get total number of steps for each day and then histogram the results:

```{r stepsTotDayPlot, fig.align='center', fig.height=6, cache=TRUE}
stepsTotDay <- with(data, tapply(steps, date, sum, na.rm=TRUE))
hist(stepsTotDay, breaks=length(stepsTotDay), main="Distribution of Total Steps Taken Per Day", xlab="Total Number of Steps per day", col="steelblue")
```

Next we will calculate and tabulate the Mean and Median of the steps taken each day:

```{r, cache=TRUE, results='asis', cache=TRUE}
## Get the mean and median values for each day
Means   <-  with(data, aggregate(x=steps, by=list(date), FUN=mean))
Medians <-  with(data, aggregate(x=steps, by=list(date), FUN=median))
## Put in a dataframe
df <- data.frame(Date= as.character(Means$Group.1), Means = Means$x, Medians = Medians$x)
## Put in a table and print it
xt <- xtable(df, caption="Mean and Median Steps Per Day", )
print(xt, type="html", floating=FALSE)
```

## What is the average daily activity pattern?

```{r timeserisplot, cache=TRUE, fig.align='center', fig.width=12}
Interval <- data[as.character(data$date)=="2012-10-01",]$interval
Interval_Avg <- with(data, aggregate(steps, list(interval), mean, na.rm=TRUE))
plot(Interval, Interval_Avg$x, "l", col="blue", main="Average Steps per 5 Minute Interval", xlab="Interval", ylab="Steps Average", lwd=2)
```

```{r maxs}
Max <- max(Interval_Avg$x)
Max_Interval <- Interval[max(Interval_Avg$x)]
cat(sprintf("Maximum average is %.2f which occurs on the %.0f 5-minute interval", Max, Max_Interval))
```

The 5-minute interval that, on average, contains the maximum number of steps is the **`r Max_Interval`** interval with the maximum average of **`r Max`** steps.

## Imputing missing values
To get the total number of missing values we use the `count` function from the `plyr` package as follows:

```{r numberofNAs}
totalNAs <- count(data[is.na(data$steps),]$steps)[2]
```

There are a total of **`r totalNAs`** rows with `NA` entries. 

The strategy we will use to fill in the missing values is to use the mean of that 5-minute interval, i.e. the values presented in the timeline plot above. 

```{r fillNAs, cache=TRUE}
## Get intervals for NAs in the data
NAIntervals <- data[is.na(data$steps),]$interval
## Subet dataframe on the NAs
Fill <- Interval_Avg[Interval_Avg$Group.1 %in% NAIntervals,]
## copy the old data set to a new one which will be modified to contain no NAs
## but first let's rearrange the columns
data <- data[c("date", "steps", "interval")]
dataNew <- data
for (i in 1:nrow(dataNew)){
        if(is.na(dataNew$steps[i])){
                dataNew$steps[i] <- Fill$x[i]                
        }
}
````

The **new data** set has no `NAs` in it as can be seen below:
```{r}
head(dataNew, 20)
```

compared to the **old data** set which still has `NAs` in it:
```{r}
head(data, 20)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r stepsTotDayPlotNew, fig.align='center', fig.height=6, cache=TRUE}
stepsTotDay <- with(dataNew, tapply(steps, date, sum, na.rm=TRUE))
hist(stepsTotDay, breaks=length(stepsTotDay), main="Distribution of Total Steps Taken Per Day", xlab="Total Number of Steps per day", col="red")
```


```{r, cache=TRUE, results='asis', cache=TRUE}
## Get the mean and median values for each day
Means   <-  with(dataNew, aggregate(x=steps, by=list(date), FUN=mean))
Medians <-  with(dataNew, aggregate(x=steps, by=list(date), FUN=median))
## Put in a dataframe
df <- data.frame(Date= as.character(Means$Group.1), Means = Means$x, Medians = Medians$x)
## Put in a table and print it
xt <- xtable(df, caption="Mean and Median Steps Per Day", )
print(xt, type="html", floating=FALSE)
```






<!--
To answer this let's first look at the total number of steps taken each day.  
For this step we will use the `lattice` package: 

```{r stepsday1, cache=TRUE}
histogram(~ steps |date, data=data)
```

We can have a more clear plot of the distributions of steps per day if we histogram the 
log of the steps as follows: 

```{r stepsday2, cache=TRUE}
histogram(~ log10(steps) |date, data=data)
```

To get the mean, median, and total number of steps taken per day we do:
-->